---
title: "Data Manipulation with `dplyr`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Once you've imported data, you're going to need to do some cleaning.

```{r}
library(tidyverse)
```

```{r}
options (tibble.print_min = 3)
```

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_df = read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```
### Select
```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
```
```{r}
select( litters_df, group:gd_of_birth)
```

```{r}
select(litters_df, -pups_survive)
```
how to remove variables from the dataset/collapse them

```{r}
select(litters_df, -pups_survive, -group)
```
how to get rid of multiple variables

```{r}
select(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
when renaming the orignial names is one the righ hand and the new name is on the left

sometimes you dont need to use select but rather rename
```{r}
rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

```{r}
select(litters_df, starts_with("gd"))
select(litters_df, ends_with ("weight"))
```
how to select certain variable in shorthand

```{r}
select(litters_df, litter_number, pups_survive, everything())
```
how to reorder the variables


### Filter
'==' is equal to and '!=' is not equal. '%in%' to detect variables that appear in a set and
'is.na()' to find missing values. 
```{r}
filter(litters_df,gd_of_birth == 20)
filter(litters_df, pups_born_alive >= 2)
filter(litters_df, pups_survive != 4)
filter (litters_df,!(pups_survive == 4))
filter(litters_df, group %in% c("Con7", "Con8"))
filter(litters_df, group == "Con7" & gd_of_birth == 20)
```
Common filtering step to omit missing data. You can use 'filter' but 'drop_na' is preferred.
drop_na(litters_data) will remove any row with a missing value
drop_na(litters_data, wt_increase) will remove rows for which wt_increase is missing.


#mutate- to create new variables
```{r}
mutate(litters_df, wt_gain= gd18_weight- gd0_weight,
       group = str_to_lower(group))
      # wt_gain_kg = wt_gain *2.2))
```
# ARRANGE- REARRGANGES THE DATA, sorts it
```{r}
head(arrange(litters_df, group, pups_born_alive), 10)
```
head function prints the first observations (in this case the first 10)

#PIPING
```{r}
litters_data_raw = read_csv("data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
```


```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(data = ., wt_gain)
```

```{r}
litters_data %>%
  lm(wt_gain ~ pups_born_alive, data = .) %>%
  broom::tidy()
```